---
title: "Part 1: SSN object development"
author: "Matt Fuller"
date: "2026-09-09"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: '3'
---

```{r setup, include=FALSE}
library(sf)
library(SSNbler)
library(SSN2)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```


Read in the necessary input data files.  

```{r}
sites_sf <- readRDS("../data/obs_df_MA_SSN.RDS")
preds_sf <- readRDS("../data/preds_df_MA_SSN.RDS") 
edges_sf <- readRDS("../data/edges_df_MA_SSN.RDS")
```

LSN build. 

```{r}
edges <-  lines_to_lsn(
  streams = edges_sf,
  lsn_path = "../ssn_obj/",
  check_topology = TRUE,
  snap_tolerance = 0.05,
  topo_tolerance = 20,
  overwrite = TRUE
)
```

Add observation sites to LSN.  

```{r}
obs <- sites_to_lsn(
  sites = sites_sf,
  edges = edges,
  lsn_path = "../ssn_obj/",
  file_name = "obs",
  snap_tolerance = 10,
  save_local = TRUE,
  overwrite = TRUE
)
```

aAdd prediction sites to LSN.  

```{r}
preds <- sites_to_lsn(
  sites = preds_sf,
  edges = edges,
  save_local = TRUE,
  lsn_path = "../ssn_obj/",
  file_name = "preds.gpkg",
  snap_tolerance = 10,
  overwrite = TRUE
)
```

Calculate upstream distance along edges.  

```{r}
edges <- updist_edges(
  edges = edges,
  save_local = TRUE,
  lsn_path = "../ssn_obj/",
  calc_length = TRUE
)

names(edges) ## View edges column names
```

Calculate upstream distance for observation and prediction sites.  

```{r}
site.list <- updist_sites(
  sites = list(
    obs = obs,
    preds = preds
  ),
  edges = edges,
  length_col = "Length",
  save_local = TRUE,
  lsn_path = "../ssn_obj/"
)

names(site.list) ## View output site.list names
```

Calculate the additive function value for tributary convergence.  

```{r}
edges <- afv_edges(
  edges = edges,
  infl_col = "TotAreaKm2",
  segpi_col = "areaPI",
  afv_col = "afvArea",
  lsn_path = "../ssn_obj/"
)
```

Calculate the additive function values for each site along the stream reaches.  

```{r}
site.list <- afv_sites(
  sites = site.list,
  edges = edges,
  afv_col = "afvArea",
  save_local = TRUE,
  lsn_path = "../ssn_obj/"
)
```


Assemble the final SSN object.  

```{r}
MA_ssn <- ssn_assemble(
  edges = edges,
  lsn_path = "../ssn_obj/",
  obs_sites = site.list$obs,
  preds_list = site.list[c("preds")],
  ssn_path = "../ssn_obj/MA_SSN.ssn",
  import = TRUE,
  check = TRUE,
  afv_col = "afvArea",
  overwrite = TRUE
)
```

Create the distance matrices for the SSN object.  

```{r}
ssn_create_distmat(MA_ssn,
                   predpts = "preds",
                   overwrite = TRUE,
                   among_predpts = TRUE)
```
