---
title: 'Part 2: Model Fit and Diagnostics'
author: "Matt Fuller"
date: "2026-09-09"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: '3'
---

# Analysis preparation  

```{r setup, include=FALSE}
library(sf)
library(SSN2)
library(tidyverse)

# load built function for running a k-fold cross validation on SSN models
source("../scripts/helperfnx/SSNkfoldcv_pbapply.R")

# load built function to add raw coef estimates/SEs to tidy() estimate table
source(file = "../scripts/helperfnx/ssn_backtransformation_fnx.R")

knitr::opts_chunk$set(echo = TRUE)
```

## Read in SSN object and distance matrices(.ssn)  

```{r}
# Read in the SSN object and prediction sites contained in the "preds" 
# predictions slot of the SSN object. (< 1 min to load)
ssn_obj <- ssn_import(path = "../ssn_obj/MA_SSN.ssn/", 
                      overwrite = TRUE, 
                      predpts = "preds")
```

## Make final adjustments to SSN object  

```{r}
# Observation site data updates
obs_df <- ssn_get_data(ssn_obj, "obs") |>
  mutate(yearf = factor(yearf),
         stn_id = factor(stn_id))

ssn_obj <- ssn_put_data(obs_df, ssn_obj, "obs")

# Prediction site data updates
preds_df <- ssn_get_data(ssn_obj, "preds") |>
  mutate(yearf = factor("0"),
         stn_id = factor(stn_id)) 

ssn_obj <- ssn_put_data(preds_df, ssn_obj, "preds")

saveRDS(ssn_obj$edges, file = "../data/ssn_obj_edges.RDS")
```

# Fit the SSN model  

```{r}
fit99 <- ssn_lm(
  formula = WT_avg ~ sAreaKm2 + sdJLtavg + sBFIc + sCPY100x + sv2elevm + sNIDDENw,
  ssn.object = ssn_obj,
  tailup_type = "mariah",
  euclid_type = "gravity",
  random = ~ stn_id + yearf,
  additive = "afvArea",
  estmethod = "reml"
)
```

# Model diagnostics and cross validation  

## Model coefficient estimate table  

```{r}
fit_esttab <- tidy(fit99) 

# grab the continuous variables from the obs data to make back transformations
cont_cov_data <- 
  select(obs_df, AreaKm2, NIDDENw, BFIc, dJLtavg, CPY100x, v2elevm) |>
  st_drop_geometry()

# generate the estimate table with raw units coefficients 
estimate_table <-
  std_to_raw_estimate_table_fnx(cont_cov_data, fit_esttab, 
                                roundval = 8, message = FALSE)

(esttab_cooling_df <- estimate_table |>  mutate(cool1C = -1/raw_est) )

write_csv(esttab_cooling_df, file = "../tables/Table2_estimate_table.csv")
```

# Cross validation  

```{r warning=FALSE}
# k-fold cross validation for the fitted SSN model to assess predictive performance
k10foldoutput <- SSNkfoldcv(fit99, kfoldn = 10, clustern = 10)
write_csv(x = k10foldoutput, file = "../tables/k10foldcv_table.csv")

# LOOCV for the fitted SSN model for comparison of diagnostic statistics to
# k-fold cross validation
fit99_loocv <- SSN2::loocv(fit99)
write_csv(x = fit99_loocv, file = "../tables/loocv_table.csv")

# Look at the variance decomposition among model structure components
fit99_varcomp <- varcomp(fit99)
write_csv(x = fit99_varcomp, file = "../tables/varcomp_table.csv")
```

# Model diagnostic plots  

```{r}
plot(fit99, which = 1)
plot(fit99, which = 2)
plot(fit99, which = 3)
plot(fit99, which = 4)
plot(fit99, which = 5)
plot(fit99, which = 6)
```

```{r}
fit99_aug <- augment(fit99)

(fit99_obspres_plot <- 
    ggplot(data = fit99_aug, aes(x = WT_avg, y = .fitted)) + 
    geom_abline(intercept = 0, slope = 1) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", formula = y ~ x) +
    theme_bw() +
    labs(title = "Observed vs Fitted", x = "Observed", y = "Fitted")
)
```


```{r fig.height = 8, fig.width=6.5}
ma_state_boundary <- 
  rnaturalearth::ne_states(country = c("United States of America"), 
                           returnclass = "sf") |>
  filter(name == "Massachusetts")

stdresiduals_neg <- ggplot() +
  # Add the state boundary
  geom_sf(data = ma_state_boundary, fill = NA, color = "grey20", linewidth = 0.5) +
  geom_sf(data = filter(fit99_aug, .std.resid <= 0) |> arrange(.std.resid), 
          aes(size = abs(.std.resid), color = .std.resid), 
          alpha = 0.3) +
  scale_color_distiller(palette = "RdBu",
    type = "div",
    direction = -1,                 # blue = low, red = high
    limits = c(-4, 4),
    name = "Standardized\nresiduals") +
  theme_minimal() +
  labs(title = "Negative standardized residuals", size = "Standardized\nresiduals")

stdresiduals_pos <- ggplot() +
  # Add the state boundary
  geom_sf(data = ma_state_boundary, fill = NA, color = "grey20", linewidth = 0.5) +
  geom_sf(data = filter(fit99_aug, .std.resid >= 0) |> arrange(desc(.std.resid)),
          aes(size = abs(.std.resid), color = .std.resid), 
          alpha = 0.3) +
  scale_color_distiller(palette = "RdBu",
    type = "div",
    direction = -1,                 # blue = low, red = high
    limits = c(-4, 4),
    name = "Standardized\nresiduals") +
  theme_minimal() +
  labs(title = "Positive standardized residuals", size = "Standardized\nresiduals")

cowplot::plot_grid(stdresiduals_pos, 
                   stdresiduals_neg + theme(legend.position = "none"),
                   nrow = 2,
                   align = "hv",
                   labels = c("A", "B"))
```



```{r}
# Compute the Torgegram
tg <- Torgegram(formula = WT_avg ~ sAreaKm2 + sdJLtavg + sBFIc + sCPY100x + 
                  sv2elevm + sNIDDENw,
                ssn.object = ssn_obj)
# Plot it
plot(tg)
```

<br>  

# Predictions  

## Recent climate scenario  

### Baseline - no restoration  

```{r}
# make predictions for scenario
augm_rcnt <- augment(x = fit99, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_rcnt, file = "../predictions/augm_rcnt.RDS")
```


### + Dam removal  

```{r}
# pull prediction data and make the update with dam removal from 'sdam1'
fit99preds_rcntdam <- ssn_get_data(fit99, name = "preds") |>
  mutate(sNIDDENw = sdam1)

# put the updated air temp data back into the model object and rename it
fit99_rcntdam <- ssn_put_data(fit99preds_rcntdam, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_rcntdam <- augment(x = fit99_rcntdam, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_rcntdam, file = "../predictions/augm_rcntdam.RDS")
```

### + Shade restoration  

```{r}
# pull prediction data and make the update with shade restoration from sCPYadd10
fit99preds_rcntshade <- ssn_get_data(fit99, name = "preds") |>
  mutate(sCPY100x = sCPYadd10)

# put the updated air temp data back into the model object and rename it
fit99_rcntshade <- ssn_put_data(fit99preds_rcntshade, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_rcntshade <- augment(x = fit99_rcntshade, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_rcntshade, file = "../predictions/augm_rcntshade.RDS")
```

### + Dam removal + shade restoration  

```{r}
# pull prediction data and make the update with dam and shade restorations
fit99preds_rcntdamshade <- ssn_get_data(fit99, name = "preds") |>
  mutate(sNIDDENw = sdam1,
         sCPY100x = sCPYadd10)


# put the updated air temp data back into the model object and rename it
fit99_rcntdamshade <- ssn_put_data(fit99preds_rcntdamshade, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_rcntdamshade <- augment(x = fit99_rcntdamshade, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_rcntdamshade, file = "../predictions/augm_rcntdamshade.RDS")
```


<br>  

## Future climate scenario  

### Baseline - no restoration  

```{r}
# Scenario: Future climate (+2degC air temp) warming                        ####
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# pull prediction data and make the update with 2C warming field
fit99preds_futr <- ssn_get_data(fit99, name = "preds") |>
  mutate(sdJLtavg = sATJL2C)

# put the updated air temp data back into the model object and rename it
fit99_futr <- ssn_put_data(fit99preds_futr, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_futr <- augment(x = fit99_futr, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_futr, file = "../predictions/augm_futr.RDS")
```


### + Dam removal  

```{r}
# pull prediction data and make the update with 2C warming and dam removal
fit99preds_futrdam <- ssn_get_data(fit99, name = "preds") |>
  mutate(sNIDDENw = sdam1,
         sdJLtavg = sATJL2C)


# put the updated air temp data back into the model object and rename it
fit99_futrdam <- ssn_put_data(fit99preds_futrdam, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_futrdam <- augment(x = fit99_futrdam, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_futrdam, file = "../predictions/augm_futrdam.RDS")
```

### + Shade restoration  

```{r}
# pull prediction data and make the update with 2C warming and shade restoration
fit99preds_futrshade <- ssn_get_data(fit99, name = "preds") |>
  mutate(sCPY100x = sCPYadd10,
         sdJLtavg = sATJL2C)


# put the updated air temp data back into the model object and rename it
fit99_futrshade <- ssn_put_data(fit99preds_futrshade, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_futrshade <- augment(x = fit99_futrshade, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_futrshade, file = "../predictions/augm_futrshade.RDS")
```

### + Dam removal + shade restoration  

```{r}
# pull prediction data and make the update with 2C warming, dam, and shade 
fit99preds_futrdamshade <- ssn_get_data(fit99, name = "preds") |>
  mutate(sNIDDENw = sdam1,
         sCPY100x = sCPYadd10,
         sdJLtavg = sATJL2C)


# put the updated air temp data back into the model object and rename it
fit99_futrdamshade <- ssn_put_data(fit99preds_futrdamshade, fit99, name = "preds")

# run predictions via augment() with the new model data
augm_futrdamshade <- augment(x = fit99_futrdamshade, newdata = "preds") |>
  left_join(select(ssn_obj$edges, rid, Shape_Leng) |> sf::st_drop_geometry(), 
            by = "rid")
saveRDS(augm_futrdamshade, file = "../predictions/augm_futrdamshade.RDS")
```

